The Astropy Project

### 2017-11-20 main.tex
address community comment

\section{Introduction} \label{sec:intro}
% first draft by Moritz (hamogu)
% heavily edited by Adrian PW - let me know if you want to revert anything!
All astronomical research makes use of software in some way.  Astronomy as a field has thus long supported the development of software tools
for astronomical tasks: from scripts that enable individual scientific research to software packages for small collaborations to data reduction pipelines for survey operations.
Some software packages are or were supported by large institutions and are intended for a wide range of users. These packages typically provide some level of documentation and user support or
training.
Other packages are developed by individual researchers or research groups and
are then typically used by smaller groups for more domain-specific purposes.
Whether for a package meant for wide distribution or for scripts and programs
for a specific research project, the implementation of astronomical software can
be eased through the use of a library that provides core functionality that is
common to many astronomical tasks.
The users of such software then also benefit from a community and ecosystem
built around a common foundation.
The \astropy project has grown to become this community, and the \astropypkg
core package a feature-rich library for \python-based astronomical software.

\inlinecomment{Tom R}{the 'this community' above is a bit too general - do we mean it is the only community in astronomy, or in the Python astronomy world? I'd change this to either 'to become this community for Python astronomy software' or 'to become such a community'. For instance yt had a community before we came along for the simulation community, so I want to make sure we don't claim we are the only community out there.}
\inlinecomment{Perry G}{Agreed}

@@

%\inlinecomment{SCO}{Table and FITS tables}

@@

% We summarize key new features or updates to \package{astropy.io} below,
% organized by sub-subpackage.

% \inlinecomment{Tom R}{is a sub-subpackage a thing, or is it just 'subpackage' all the way down?}

\subsubsection{ASCII}
% \subsubsection{Enhanced CSV format}

One of the problems when storing a table in an ASCII format is
preserving table meta-data such as comments, keywords and column data
types, units, and descriptions. The newly defined \emph{Enhanced
Character Separated Values} (ECSV,  \citet{ape6}) format makes it
possible to write a table to an ASCII-format file and read it back
with no loss of information. The ECSV format has been designed to be
both human-readable and compatible with most simple CSV readers.

% \subsubsection{\texttt{.ascii}: fast readers and writers}

The \package{astropy.io.ascii} subpackage now includes a significantly faster
Cython/C engine for reading and writing ASCII files. This is available for the
most common formats.  On average the new engine is about 4 to 5 times faster
than the corresponding pure-\python implementation, and is often comparable to
the speed of the \package{Pandas} \citep{pandas} ASCII file interface.  The
fast reader has parallel processing option that allows harnessing multiple
cores for input parsing to achieve even greater speed gains.  By default,
\texttt{read()} and \texttt{write()} will attempt to use the fast C engine
when dealing with compatible formats. Certain features of the full read
/ write interface are not available in the fast version, in which case the
pure-\python version will automatically be used.

% \subsubsection{\texttt{.ascii}: HTML tables}

The \package{astropy.io.ascii} subpackage now provides the capability
to read a table within an HTML file or web URL into an astropy
\texttt{Table} object. Conversely a \texttt{Table} object can now
be written out as an HTML table.

\subsubsection{FITS}

The \package{astropy.io.fits} subpackage started as a direct port of the
PyFITS project \citep{PyFITS}. Therefore it is pretty stable, with mostly bug
fixes but also a few new features and performance improvements.  The API
remains compatible with PyFITS, which is now deprecated in favor of
\astropypkg.

Command-line scripts are now available for printing a summary of the HDUs in
a FITS file(s) (\texttt{fitsinfo}) and for printing the header information to
the screen in a human-readable format (\texttt{fitsheader}).

FITS files are now loaded \emph{lazily} (the default behavior, which can be
deactivated if needed), where all HDUs are not loaded until they are
requested. This should provide substantial speedups for situations using the
convenience functions (e.g., \texttt{getheader()} or \texttt{getdata()}) to
get HDU’s that are near the front of a file with many HDUs.

% \inlinecomment{Tom R}{too detailed? I don't think we need to mention every single performance enhancement, right?}

% \subsubsection{\texttt{.misc}: YAML serialization}

% The new \package{astropy.io.misc.yaml} module allows converting astropy
% objects into a standard YAML format.  This can be used beyond
% \texttt{astropy.io.misc.yaml}: for example, to serialize the metadata of
% tables before saving to other formats like HDF5.

>>>>

\section{Introduction} \label{sec:intro}
% first draft by Moritz (hamogu)
% heavily edited by Adrian PW - let me know if you want to revert anything!
All astronomical research makes use of software in some way.  Astronomy as a field has thus long supported the development of software tools
for astronomical tasks: from scripts that enable individual scientific research to software packages for small collaborations to data reduction pipelines for survey operations.
Some software packages are or were supported by large institutions and are intended for a wide range of users. These packages typically provide some level of documentation and user support or
training.
Other packages are developed by individual researchers or research groups and
are then typically used by smaller groups for more domain-specific purposes.
Whether for a package meant for wide distribution or for scripts and programs
for a specific research project, the implementation of astronomical software can
be eased through the use of a library that provides core functionality that is
common to many astronomical tasks.
The users of such software then also benefit from a community and ecosystem
built around a common foundation.
The \astropy project has grown to become this community for \python astronomy
software, and the \astropypkg core package a feature-rich library for
\python-based astronomical software.

%\inlinecomment{Tom R}{the 'this community' above is a bit too general - do we mean it is the only community in astronomy, or in the Python astronomy world? I'd change this to either 'to become this community for Python astronomy software' or 'to become such a community'. For instance yt had a community before we came along for the simulation community, so I want to make sure we don't claim we are the only community out there.}
%\inlinecomment{Perry G}{Agreed}

@@

%\inlinecomment{SCO}{Pyfits deprecation in favor of io.fits ?}
%\inlinecomment{SCO}{Table and FITS tables}

@@

We summarize key new features or updates to \package{astropy.io} below,
organized by sub-subpackage.

\inlinecomment{Tom R}{is a sub-subpackage a thing, or is it just 'subpackage' all the way down?}

\subsubsection{Enhanced CSV format}

One of the problems when storing a table in an ASCII format is
preserving table meta-data such as comments, keywords and column data
types, units, and descriptions. Using the newly defined \emph{Enhanced
Character Separated Values} (ECSV,  \citet{ape6}) format it is now
possible to write a table to an ASCII-format file and read it back
with no loss of information. The ECSV format has been designed to be
both human-readable and compatible with most simple CSV readers.

\subsubsection{\texttt{.ascii}: fast readers and writers}

The \package{astropy.io.ascii} module now includes a significantly
faster Cython/C engine for reading and writing ASCII files. This is
available for the most common formats.

%following formats: \texttt{basic},
%      \texttt{commented\_header}, \texttt{csv}, \texttt{no\_header},
%      \texttt{rdb}, and \texttt{tab}.
On average the new engine is about
4 to 5 times faster than the corresponding pure-\python
implementation, and is often comparable to the speed of the
\package{Pandas} \citep{pandas} ASCII file interface.  The fast
reader has parallel processing option that allows harnessing
multiple cores for input parsing to achieve even greater speed
gains.

By default, \texttt{read()} and \texttt{write()} will attempt to use
the fast C engine when dealing with compatible formats. Certain
features of the full read / write interface are not available in the
fast version, in which case the pure-\python version will automatically
be used.

\subsubsection{\texttt{.ascii}: HTML tables}

The \package{astropy.io.ascii} subpackage now provides the capability
to read a table within an HTML file or web URL into an astropy
\texttt{Table} object. This requires the \package{BeautifulSoup4}
package to be installed.  Conversely a \texttt{Table} object can now
be written out as an HTML table.

\subsubsection{\texttt{.fits}: command-line scripts (\texttt{fitsheader}, \texttt{fitsinfo})}

The \package{astropy.io.fits} subpackage now provides command-line
scripts for printing a summary of the HDUs in a FITS file(s)
(\texttt{fitsinfo file.fits}) and for printing the header information
to the screen in a human-readable format (\texttt{fitsheader
file.fits}).

\subsubsection{\texttt{.fits}: lazy loading }

The \package{astropy.io.fits} subpackage now supports \emph{lazy
loading}, where all HDUs are not loaded until they are requested (or
the file is closed). This should provide substantial speedups for
situations using the convenience functions (e.g., \texttt{getheader()}
or \texttt{getdata()}) to get HDU’s that are near the front of a file
with many HDUs.

\inlinecomment{Tom R}{too detailed? I don't think we need to mention every single performance enhancement, right?}

\subsubsection{\texttt{.misc}: YAML serialization}

The new \package{astropy.io.misc.yaml} module allows converting
astropy objects into a standard YAML format.
This can be used beyond \texttt{astropy.io.misc.yaml}: for example, to
serialize the metadata of tables before saving to other formats like
HDF5.
